Who does big team science?

### 2023-11-01 manuscript_scopus.tex
updating the manuscript

The credibility movement was originally defined by a focus on large
scale replications used in collaborative environments\textsuperscript{11}.
Generally, the movement appears to have been driven by early career
researchers (i.e., those who are within five years of their first
appointment)\textsuperscript{12}; however, there are no large meta-scientific
investigations on this specific topic to date. Potentially, the lack of
investigation is tied to the newness of the large-scale research in many
fields, as it is only in recent years that publications like the Open
Science Collaboration\textsuperscript{13}, Many Labs
Collaborations\textsuperscript{14--20} or the first papers from the
Psychological Science Accelerator\textsuperscript{21--27}. Generally, the researcher incentive for
replication was low for three reasons. First, journals often prioritize
``novel'' or new results which led to rejection of replication manuscripts
and publication bias\textsuperscript{28--30}. Second,
the ``failure'' to replicate was often placed on the replication team as
``bad science'' rather than a careful consideration of publication biases
and (potential) questionable research practices\textsuperscript{5,18}. Last, why should someone want to spend time and resources
on an answer we already ``know''\textsuperscript{31,32}?

However, the success and interest in the large-scale reproducibility
projects\textsuperscript{13,33}, paired
with the meta-scientific publications focusing on researcher practices
and incentive structures\textsuperscript{34,35} led to a
change in journal guidelines and incentives for researchers interested
in participating in large-scale replication studies\textsuperscript{36--39}. For example, the support
for Registered Reports, papers accepted before the data has been
collected\textsuperscript{40,41}, and entire sub-sections of
journals devoted to only replication studies (e.g., \emph{Nature, Royal
Society Open Science, Advances in Methods and Practices in Psychological
Science}) has allowed researchers to invest in projects that they know
should be published when the project is complete. Further, the
implementation of the Transparency and Openness Guidelines\textsuperscript{38}
and the Contributor Role Taxonomy (CRediT) system\textsuperscript{42} have
pushed journals and researchers to promote more open, inclusive
publication practices.

The credibility movement has been mirrored by the calls for
diversification or de-WEIRDing (e.g., Western, Educated, Industrialized,
Rich, and Democratic) scientific research\textsuperscript{43--45} by improving representation in research samples. Like the
large-scale studies in Physics\textsuperscript{46,47} and
Biology\textsuperscript{48}, the Social Sciences struggle to represent the
breadth of humanity across both researcher and population
characteristics. Now, grassroots organizations, such as the
Psychological Science Accelerator\textsuperscript{26}, ManyBabies
(\url{https://manybabies.github.io/}), NutNet (\url{https://nutnet.org/}), and
DRAGNet (\url{https://dragnetglobal.weebly.com/}) can begin to tackle these
issues by recruiting research labs from all over the globe to provide
diversity in geographic, linguistic, and researcher representation.
Publications have examined the global understanding of morality, face
processing, COVID-19 information signaling, and more\textsuperscript{21,23--25,27,49}. While these organizations and one-time groups
for BTS studies have provided an incredible wealth of data for the
scientific community, we do not yet know exactly \emph{who} is involved with,
and benefits from, the BTS and credibility movement. Publications on BTS
generally explore challenges, lessons learned, and the need for BTS\textsuperscript{2,3}.

Therefore, the goal of this manuscript is to examine the \emph{people}
involved in BTS projects. We specifically examined the themes of
inclusivity, research careers, and research globalization. We see an
increasing interest and number of publications in BTS but we do not yet
know if this uptick in large-scale projects has diversified the \emph{people}
involved in BTS. While a few publications have noted that BTS appears to
be early career researchers\textsuperscript{12}, no one has systematically
investigated this perception. Further, it is unclear if the focus of
de-WEIRDing science has only focused on the representation of the
research participants or if it has also improved the representation of
researchers outside of North America and Europe. Last, who runs these
BTS projects? Do we see an increase in diversity for the authors who
generally receive the most credit for these projects (i.e., first
several author(s) and last author)? As hiring and promoting practices
often place a heavy weight on publications and especially ``influential''
publications, it becomes necessary to critically examine the
representation present in authorship in BTS projects.

@@

\begin{itemize}
\tightlist
\item
  Research Question 1: What publication sources publish big team
  science papers?
\item
  Research Question 2: What are the types of articles that are being
  published in big team science?
\item
  Research Question 3: Who is involved in big team science?
\end{itemize}

@@

We have defined BTS publications as publications with at least ten
authors at ten different institutions that were published in
peer-reviewed journals or had posted a full paper pre-print. We used
data from 1970 and forward in the Scopus database, as it is noted online
that this time period includes cited references for calculation of
several of our variables described below. We will analyze our results
based on four subject areas present in the Scopus database: Physical
Sciences, Health Sciences, Social Sciences, and Life Sciences. We
filtered the database to include articles, articles in press, business
articles, conference papers, data papers, preprints, and surveys using
Elsevier's classification system. This project was supported by access
to the Scopus database through the International Center for the Study of
Research.

@@

\hypertarget{rq3-author-information}{%
\subsubsection{RQ3: Author Information}\label{rq3-author-information}}

@@

\textbf{\emph{Career Length}}. Career length for each author was defined as the
year of the first publication minus the current year listed for each
author.

\textbf{\emph{Institution and Geopolitical Region}}. We used the affiliation ids
and country to gather information about the places of education and/or
employment for authors. Geopolitical region was created by binning these
codes into United Nation Regions.

@@

\textbf{\emph{Publication Metrics}}. For each author, we calculated the total
number of publications, and the h-index. The h-index represents the
highest \emph{h} number of publications that have at least \emph{h} citations.
\emph{h}-count was only used for descriptive statistics.

@@

\hypertarget{rq3-authors.}{%
\subsection{RQ3: Authors.}\label{rq3-authors.}}

@@

\textbf{\emph{Career Length}}.

Figure \ref{fig:fig-career} portrays the average career length for
authors involved in BTS publications across years. Career length was
defined as the year of first publication minus the current year, and
higher numbers mean longer careers. To analyze trends over time, we
calculated the average career length for each publication (i.e., average
author career lengths to create one score for each paper) and analyzed a
regression analysis using career length to predict year of publication.
In order to show variance between individuals, we calculated the
standard deviation of career length for each publication and used this
variance as an additional predictor.

@@

\begin{figure}
\centering
\includegraphics{manuscript_scopus_files/figure-latex/fig-heatmap-1.pdf}
\caption{\label{fig:fig-heatmap}Heatmap results of regression analyses for career length, number of publications, and geopolitical diversity within the region. The top figure represents all results together for comparison across analyses. The bottom row represents individual heatmaps for each hypothesis to distinguish small differences between subject areas for those research questions. Non-significant results are indicated with NS on the plot.}
\end{figure}

\textbf{\emph{Institution}}.

The total number of unique affiliation across all papers was 463876.

\textbf{\emph{Publication Metrics}}.

The average number of publications by authors on big team science papers
is \emph{M} = 38.37 (\emph{SD} = 102.54). The
publication counts were averaged across authors for each publication,
and then these average publication counts were averaged across
publications \emph{M} = 162.50 (\emph{SD} =
155.17). The average variability (i.e., the average
standard deviation with authors of a manuscript) with publication counts
of a paper was \(M_{SD}\) = 164.27 (\(SD_{SD}\) =
127.21).

We used the same analyses described in the career length section to
analyze trends over time. An increasing slope over time indicates that
individuals who are publishing more are more represented in BTS over
time (i.e., increasing numbers of scholars with higher publication
rates), while a negative slope indicates more researchers with less
publications. A positive slope for the standard deviation of publication
metrics indicates increasing variance over time (i.e., more diversity in
the individual publication rates), while a negative slope would indicate
less diversity in researchers over time. While publication rates do not
represent value as a researcher, they are often used in hiring and
promotion decisions, and we used this variable as a proxy to gauge the
diversity in scholars represented in big teams. As shown in Figure
\ref{fig:fig-heatmap} publication metrics were generally negative for
the average publication metrics, indicating more scholars over time with
lower numbers of publications with the strongest effects in Health and
Social Sciences. The variability of publication counts was not
significant for the Life Sciences but was negative for the Physical
Sciences (less variability over time) and positive for Social and Health
Sciences (more variability and over time). This result indicates that
the Physical Sciences are trending toward scholars with less
publications but also less diverse in number of publications, while the
Health and Social Sciences see more diversity in publication counts and
less published scholars overall.

@@

\begin{figure}
\centering
\includegraphics{manuscript_scopus_files/figure-latex/fig-map-both-1.pdf}
\caption{\label{fig:fig-map-both}Geopolitical regions represented in big-team science publications versus all publications. The mosaic plot is grouped by UN subregion with the largest number of publications on the left and smallest on the top right.}
\end{figure}

@@

In this investigation, we explored the publication rates, areas, and
researchers involved in big team science publications. Over a
half-million articles were published in nearly 15,000 journals since
1970 that qualified as big team science articles (at least ten authors
and ten different affiliations). The areas of publication were aligned
to cancer and genetics research in medicine and oncology for Health and
Life Sciences, physics and chemistry for the Physical Sciences, and
psychology for the Social Sciences All areas of research show an
explosion in growth in the number of publications and the number of
authors included on manuscripts, replicating previous investigations
into this topic area\textsuperscript{50--52}.

Our investigation expands into an exploration of the researchers who
publish in big teams focusing on career length, publication metrics, and
geopolitical affiliation. The number of earlier career scholars is
increasing in publications across the years, indicating that big teams
may be more accessible to different types of individuals, not just
older, more established researchers. This result is especially
interesting given the publish-or-perish model still present in most
institutions, as it may seem that large-scale projects could be a risky
choice for non-permanent researchers. In the authors' experience, big
team projects are often quite slow to publication, incentives may be low
for non corresponding authors if institutions do not value papers
without lead authorship, and there is no guarantee for publication even
with a large group. However, with a large team the distribution of work
could imply less effort on individual non-leading members, and research
has shown that larger-team publications do receive more citations and
appear to have higher impact\textsuperscript{53}.

The results for the number of publications by big team researchers
mimics the findings from career length, with a smaller effect size. In
general, it appears that there is a decrease in the average number of
publications a researcher has when publishing on a big team science
paper over time. This result is likely attributable to the number of
early career scholars joining projects, but also may support increased
accessibility for individuals to be involved in this type of research.
Globalization, the internet, and the focus on interdisciplinary research
are potentially driving forces behind our results, but, hopefully, the
results also point to a decline in scientific gatekeeping\textsuperscript{54,55}.

The variability in the types of researchers involved in publications
also decreased across time in most areas of science with a decrease in
variability for career length. As mentioned, an increase in early career
researchers and numbers of publications could explain this effect
mathematically, potentially with other social influences mentioned
above. The variability in the number of publications is decreasing in
the Physical Sciences, mirroring the career length results, but the
opposite effect was found in the Health and Social Sciences. We see no
clear reason why career variability would decrease while the variability
in the number of publications would increase. The effect sizes for
career length were much larger than the effects for number of
publications. One speculation is the increasing requirements for a
competitive faculty role application. Given the limited number of
positions, one potential way to distinguish their application would be a
larger number of publications in their early career\textsuperscript{56,57}.

The number of geopolitical entities for researcher affiliation is
increasing over time, showing the results of globalization and the
ability to connect across time zones and cultures\textsuperscript{58}. While our
definition of big team science required at least ten different
institutional affiliations, we did not filter papers by geopolitical
region, and thus, a manuscript could rely solely on institutions within
a single country. The Physical Sciences did not show an increase in
diversity of regions represented, however, it could be argued that the
development of large research centers like CERN forced earlier diversity
than other sciences (i.e., because CERN specifically recruited
scientists from sponsoring nations). The Life, Health, and Social
Sciences saw an increase in the number of regions represented with the
highest increase in the Social Sciences. This result likely corresponds
with an increased interest in big team science publications in
psychology\textsuperscript{2,3}, and the desire to diversify the
populations represented in psychological research\textsuperscript{43,44}.

While publications on the whole are diversifying, we did not yet find
equality in the representation for first/corresponding author spots
versus all other authors. In general, first authors appear to be less
diverse, representing more European and North American authors, while
other authors include more Asian and African authors. These effect sizes
were often small, but the inequality still persists across years, even
if they are slowly decreasing. Diverse teams are more likely to have
papers with stronger ``impact''\textsuperscript{59--62} with higher citation metrics for more diverse author lists.
The introduction of contributorship models (e.g., CRediT\textsuperscript{42})
will hopefully continue to push these effects down, as they highlight
each individual's contribution to a manuscript.

@@

\leavevmode\vadjust pre{\hypertarget{ref-sinatra2015}{}}%
\CSLLeftMargin{50. }%
\CSLRightInline{Sinatra, R., Deville, P., Szell, M., Wang, D. \& Barabási, A.-L. \href{https://doi.org/10.1038/nphys3494}{A century of physics}. \emph{Nature Physics} \textbf{11}, 791--796 (2015).}

\leavevmode\vadjust pre{\hypertarget{ref-wuchty2007}{}}%
\CSLLeftMargin{51. }%
\CSLRightInline{Wuchty, S., Jones, B. F. \& Uzzi, B. \href{https://doi.org/10.1126/science.1136099}{The increasing dominance of teams in production of knowledge}. \emph{Science} \textbf{316}, 1036--1039 (2007).}

\leavevmode\vadjust pre{\hypertarget{ref-hunter2008}{}}%
\CSLLeftMargin{52. }%
\CSLRightInline{Hunter, L. \& Leahey, E. \href{https://doi.org/10.1007/s12108-008-9042-1}{Collaborative Research in Sociology: Trends and Contributing Factors}. \emph{The American Sociologist} \textbf{39}, 290--306 (2008).}

\leavevmode\vadjust pre{\hypertarget{ref-lariviuxe8re2015}{}}%
\CSLLeftMargin{53. }%
\CSLRightInline{Larivière, V., Gingras, Y., Sugimoto, C. R. \& Tsou, A. \href{https://doi.org/10.1002/asi.23266}{Team size matters: Collaboration and scientific impact since 1900}. \emph{Journal of the Association for Information Science and Technology} \textbf{66}, 1323--1332 (2015).}

\leavevmode\vadjust pre{\hypertarget{ref-lu2007}{}}%
\CSLLeftMargin{54. }%
\CSLRightInline{Lu, Y. \href{https://doi.org/10.1016/j.lisr.2006.10.007}{The human in human information acquisition: Understanding gatekeeping and proposing new directions in scholarship}. \emph{Library \& Information Science Research} \textbf{29}, 103--123 (2007).}

\leavevmode\vadjust pre{\hypertarget{ref-siler2015}{}}%
\CSLLeftMargin{55. }%
\CSLRightInline{Siler, K., Lee, K. \& Bero, L. \href{https://doi.org/10.1073/pnas.1418218112}{Measuring the effectiveness of scientific gatekeeping}. \emph{Proceedings of the National Academy of Sciences} \textbf{112}, 360--365 (2015).}

\leavevmode\vadjust pre{\hypertarget{ref-caplow2017}{}}%
\CSLLeftMargin{56. }%
\CSLRightInline{Caplow, T. \emph{The academic marketplace}. (Routledge, 2017). doi:\href{https://doi.org/10.4324/9781351305969}{10.4324/9781351305969}.}

\leavevmode\vadjust pre{\hypertarget{ref-kyvik2003}{}}%
\CSLLeftMargin{57. }%
\CSLRightInline{Kyvik, S. \href{https://doi.org/10.1023/A:1025475423482}{Changing trends in publishing behaviour among university faculty, 1980-2000}. \emph{Scientometrics} \textbf{58}, 35--48 (2003).}

\leavevmode\vadjust pre{\hypertarget{ref-xie2014}{}}%
\CSLLeftMargin{58. }%
\CSLRightInline{Xie, Y. \href{https://doi.org/10.1126/science.1252743}{{``}Undemocracy{''}: Inequalities in science}. \emph{Science} \textbf{344}, 809--810 (2014).}

\leavevmode\vadjust pre{\hypertarget{ref-freeman2015}{}}%
\CSLLeftMargin{59. }%
\CSLRightInline{Freeman, R. B. \& Huang, W. \href{https://doi.org/10.1086/678973}{Collaborating with People Like Me: Ethnic Coauthorship within the United States}. \emph{Journal of Labor Economics} \textbf{33}, S289--S318 (2015).}

\leavevmode\vadjust pre{\hypertarget{ref-hinnant2012}{}}%
\CSLLeftMargin{60. }%
\CSLRightInline{Hinnant, C. C. \emph{et al.} \href{https://doi.org/10.1016/j.lisr.2012.03.001}{Author-team diversity and the impact of scientific publications: Evidence from physics research at a national science lab}. \emph{Library \& Information Science Research} \textbf{34}, 249--257 (2012).}

\leavevmode\vadjust pre{\hypertarget{ref-jones2008}{}}%
\CSLLeftMargin{61. }%
\CSLRightInline{Jones, B. F., Wuchty, S. \& Uzzi, B. \href{https://doi.org/10.1126/science.1158357}{Multi-university research teams: Shifting impact, geography, and stratification in science}. \emph{Science} \textbf{322}, 1259--1262 (2008).}

>>>>

The credibility movement was originally defined by a focus on large
scale replications used in collaborative environments\textsuperscript{11}.
Generally, the movement appears to have been driven by early career
researchers (i.e., those who are within five years of their first
appointment)\textsuperscript{12}; however, there are no large meta-scientific
investigations on this specific topic to date. Potentially, the lack of
investigation is tied to the newness of the large-scale research in many
fields, as it is only in recent years that publications like the Open
Science Collaboration\textsuperscript{13}, Many Labs
Collaborations\textsuperscript{14--20} or the first papers from the
Psychological Science Accelerator\textsuperscript{21--27}. Generally, the researcher incentive for
replication and/or involvement in big-team projects was low for three reasons. First, journals often prioritize
``novel'' or new results which led to rejection of replication manuscripts
and publication bias\textsuperscript{28--30}. Second,
the ``failure'' to replicate was often placed on the replication team as
``bad science'' rather than a careful consideration of publication biases
and (potential) questionable research practices\textsuperscript{5,18}. Last, why should someone want to spend time and resources
on an answer we already ``know''\textsuperscript{31,32}?

However, the success and interest in the large-scale reproducibility
projects\textsuperscript{13,33}, paired
with the meta-scientific publications focusing on researcher practices
and incentive structures\textsuperscript{34,35} led to a
change in journal guidelines and incentives for researchers interested
in participating in large-scale replication studies\textsuperscript{36--39}. In some fields, the
replication movement demonstrated that large-scale teams were a
practical (and publishable) solution to answering research questions
in generalizable way. The support
for Registered Reports, papers accepted before the data has been
collected\textsuperscript{40,41}, has allowed researchers to
invest in projects that they know
should be published when the project is complete. Further, the
implementation of the Transparency and Openness Guidelines\textsuperscript{38}
and the Contributor Role Taxonomy (CRediT) system\textsuperscript{42} have
pushed journals and researchers to promote more open, inclusive
publication practices.

Beyond the replication movement, the credibility movement has\\
mirrored calls for
diversification or de-WEIRDing (e.g., Western, Educated, Industrialized,
Rich, and Democratic) scientific research\textsuperscript{43--45} by improving representation in research samples. Like the
large-scale studies in Physics\textsuperscript{46,47} and
Biology\textsuperscript{48}, the Social Sciences struggle to represent the
breadth of humanity across both researcher and population
characteristics. Now, grassroots organizations, such as the
Psychological Science Accelerator\textsuperscript{26}, ManyBabies
(\url{https://manybabies.github.io/}), NutNet (\url{https://nutnet.org/}), and
DRAGNet (\url{https://dragnetglobal.weebly.com/}) can begin to tackle these
issues by recruiting research labs from all over the globe to provide
diversity in geographic, linguistic, and researcher representation.
Publications have examined the global understanding of morality, face
processing, COVID-19 information signaling, and more\textsuperscript{21,23--25,27,49}. While these organizations and one-time groups
for BTS studies have provided an incredible wealth of data for the
scientific community, we do not yet know exactly \emph{who} is involved with,
and benefits from, the BTS and credibility movement. Publications on BTS
generally explore challenges, lessons learned, and the need for BTS\textsuperscript{2,3}.

Therefore, the goal of this manuscript is to examine both the \emph{publications} and \emph{people} involved in BTS projects. We present descriptive information about the publication sources and types of articles that we classified as BTS projects to demonstrate what areas of research show large-scale research. Next, we examine the individuals involved in those projects for descriptive and predictive purposes. To describe the people involved in BTS projects, we used education, types of publications (i.e., articles, preprints, books, etc.) from BTS individuals, and publication metrics.

For predictive statistics, we explored the change in diversity of authors over time. We see an increasing interest and number of publications in BTS but we do not yet
know if this uptick in large-scale projects has diversified the \emph{people}
involved in BTS. While a few publications have noted that BTS appears to
be early career researchers\textsuperscript{12}, no one has systematically
investigated this perception. Further, it is unclear if the focus of
de-WEIRDing science has only focused on the representation of the
research participants or if it has also improved the representation of
researchers outside of North America and Europe. Last, who runs these
BTS projects? Do we see an increase in diversity for the authors who
generally receive the most credit for these projects (i.e., first
several author(s) and last author)? As hiring and promoting practices
often place a heavy weight on publications and especially ``influential''
publications, it becomes necessary to critically examine the
representation present in authorship in BTS projects.

@@

\begin{itemize}
\tightlist
\item
  Research Question 1: What publication sources publish big team
  science papers?
\item
  Research Question 2: What are the types of articles that are being
  published in big team science?
\item
  Research Question 3: Who is involved in big team science?
\item
  Research Question 4: How has the diversity of those involved in big team science changed over time?
\end{itemize}

@@

We have defined BTS publications as publications with at least ten
authors at ten different institutions that were published in
peer-reviewed journals or had posted a full paper pre-print. While this
definition is a somewhat arbitrary choice, we separate this research
from research on team science that uses any multi-university collaboration
as a definition\textsuperscript{50} to focus on larger sized teams rather
than teams of any size. With at least ten institutions, the complexities
of infrastructure, resources, tenure and promotion policies, ethics review, and more can occur\textsuperscript{3}. Therefore, we believe this choice selects publications that would be ``big'' teams and those potential obstacles.

We used
data from 1970 and forward in the Scopus database, as it is noted online
that this time period includes cited references for calculation of
several of our variables described below. We will analyze our results
based on four subject areas present in the Scopus database: Physical
Sciences, Health Sciences, Social Sciences, and Life Sciences. We
filtered the database to include articles, articles in press, business
articles, conference papers, data papers, preprints, and surveys using
Elsevier's classification system. This project was supported by access
to the Scopus database through the International Center for the Study of
Research.

@@

\hypertarget{rq3-author-descriptive-statistics}{%
\subsubsection{RQ3: Author Descriptive Statistics}\label{rq3-author-descriptive-statistics}}

@@

@@

\textbf{\emph{Publication Metrics}}. For each author, we calculated the number of publications and the h-index. The h-index represents the highest \emph{h} number of publications that have at least \emph{h} citations.

\textbf{\emph{Institutions}}. We report the number of institutions involved in big team science publications.

\hypertarget{rq4-author-diversity-statistics}{%
\subsubsection{RQ4: Author Diversity Statistics ~}\label{rq4-author-diversity-statistics}}

\textbf{\emph{Seniority}}. Career length for each author was defined as the
year of the first publication minus the current year listed for each
author. Number of publications included the number of unique entries an author was included in the database. Career length and number of publications was used as a proxy for the ``age'' or ``seniority'' of a scholar to assess if only senior scholars were

\textbf{\emph{Geopolitical Region}}. Geopolitical region was created by binning country code identifiers into the 17 identified United Nation Regions.

@@

\hypertarget{rq3-author-descriptive-statistics-1}{%
\subsection{RQ3: Author Descriptive Statistics}\label{rq3-author-descriptive-statistics-1}}

@@

\textbf{\emph{Publication Metrics}}. The average number of publications by authors on big team science papers
is \emph{M} = 38.37 (\emph{SD} = 102.54). The
publication counts were averaged across authors for each publication,
and then these average publication counts were averaged across
publications \emph{M} = 162.50 (\emph{SD} =
155.17). The average variability (i.e., the average
standard deviation with authors of a manuscript) with publication counts
of a paper was \(M_{SD}\) = 164.27 (\(SD_{SD}\) =
127.21).

\textbf{\emph{Institutions}}. The total number of unique affiliation across all papers was 463876.

\hypertarget{rq4-author-diversity-statistics-1}{%
\subsection{RQ4: Author Diversity Statistics}\label{rq4-author-diversity-statistics-1}}

\textbf{\emph{Seniority}}. Figure \ref{fig:fig-career} portrays the average career length for
authors involved in BTS publications across years. Career length was
defined as the year of first publication minus the current year, and
higher numbers mean longer careers. To analyze trends over time, we
calculated the average career length for each publication (i.e., average
author career lengths to create one score for each paper) and analyzed a
regression analysis using career length to predict year of publication.
In order to show variance between individuals, we calculated the
standard deviation of career length for each publication and used this
variance as an additional predictor.

@@

\begin{figure}
\centering
\includegraphics{manuscript_scopus_files/figure-latex/fig-heatmap-1.pdf}
\caption{\label{fig:fig-heatmap}Heatmap results of regression analyses for career length, number of publications, and geopolitical diversity within the region. Each square represents a \emph{b} value or the slope of the predictor (x-axis) onto the dependent variable (each panel). Slopes included both the overall value of the predictor (\(b\), \(b_M\)) and the standard deviation of the predictor over time (\(b_{SD}\)). The color of the square represents the strength of the predictor. The top figure represents all results together for comparison across analyses. The bottom row represents individual heatmaps for each hypothesis to distinguish small differences between subject areas for those research questions. Non-significant results are indicated with NS on the plot.}
\end{figure}

We used the same analyses using number of publications to represent diversity instead of career length. An increasing slope over time indicates that
individuals who are publishing more are more represented in BTS over
time (i.e., increasing numbers of scholars with higher publication
rates), while a negative slope indicates more researchers with less
publications. A positive slope for the standard deviation of publication
metrics indicates increasing variance over time (i.e., more diversity in
the individual publication rates), while a negative slope would indicate
less diversity in researchers over time. While publication rates do not
represent value as a researcher, they are often used in hiring and
promotion decisions, and we used this variable as a proxy to gauge the
diversity in scholars represented in big teams. As shown in Figure
\ref{fig:fig-heatmap} publication metrics were generally negative for
the average publication metrics, indicating more scholars over time with
lower numbers of publications with the strongest effects in Health and
Social Sciences. The variability of publication counts was not
significant for the Life Sciences but was negative for the Physical
Sciences (less variability over time) and positive for Social and Health
Sciences (more variability and over time). This result indicates that
the Physical Sciences are trending toward scholars with less
publications but also less diverse in number of publications, while the
Health and Social Sciences see more diversity in publication counts and
less published scholars overall.

@@

\begin{figure}
\centering
\includegraphics{manuscript_scopus_files/figure-latex/fig-map-both-1.pdf}
\caption{\label{fig:fig-map-both}Geopolitical regions represented in big-team science publications versus all publications. The mosaic plot is grouped by UN subregion with the largest number of publications starting on the bottom left and smallest on the top right. Therefore, North America represents the largest number of authors within BTS (i.e., bottom right, then separated into the geopolitical areas within that subregion), followed by Eastern Europe (top left), and so on.}
\end{figure}

@@

In this investigation, we explored the publication rates, areas, and
researchers involved in big team science publications. Over a
half-million articles were published in nearly 15,000 journals since
1970 that qualified as big team science articles (at least ten authors
and ten different affiliations). The areas of publication were aligned
to cancer and genetics research in medicine and oncology for Health and
Life Sciences, physics and chemistry for the Physical Sciences, and
psychology for the Social Sciences All areas of research show an
explosion in growth in the number of publications and the number of
authors included on manuscripts, replicating previous investigations
into this topic area\textsuperscript{51--53}.

Our investigation expands into an exploration of the researchers who
publish in big teams focusing on career length, publication metrics, and
geopolitical affiliation. The number of earlier career scholars is
increasing in publications across the years, indicating that big teams
may be more accessible to different types of individuals, not just
older, more established researchers. This result is especially
interesting given the publish-or-perish model still present in most
institutions, as it may seem that large-scale projects could be a risky
choice for non-permanent researchers. In the authors' experience, big
team projects are often quite slow to publication, incentives may be low
for non corresponding authors if institutions do not value papers
without lead authorship, and there is no guarantee for publication even
with a large group. However, with a large team the distribution of work
could imply less effort on individual non-leading members, and research
has shown that larger-team publications do receive more citations and
appear to have higher impact\textsuperscript{54}.

The results for the number of publications by big team researchers
mimics the findings from career length, with a smaller effect size. In
general, it appears that there is a decrease in the average number of
publications a researcher has when publishing on a big team science
paper over time. This result is likely attributable to the number of
early career scholars joining projects, but also may support increased
accessibility for individuals to be involved in this type of research.
Globalization, the internet, and the focus on interdisciplinary research
are potentially driving forces behind our results, but, hopefully, the
results also point to a decline in scientific gatekeeping\textsuperscript{55,56}.

The variability in the types of researchers involved in publications
also decreased across time in most areas of science with a decrease in
variability for career length. As mentioned, an increase in early career
researchers and numbers of publications could explain this effect
mathematically, potentially with other social influences mentioned
above. The variability in the number of publications is decreasing in
the Physical Sciences, mirroring the career length results, but the
opposite effect was found in the Health and Social Sciences. We see no
clear reason why career variability would decrease while the variability
in the number of publications would increase. The effect sizes for
career length were much larger than the effects for number of
publications. One speculation is the increasing requirements for a
competitive faculty role application. Given the limited number of
positions, one potential way to distinguish their application would be a
larger number of publications in their early career\textsuperscript{57,58}.

The number of geopolitical entities for researcher affiliation is
increasing over time, showing the results of globalization and the
ability to connect across time zones and cultures\textsuperscript{59}. While our
definition of big team science required at least ten different
institutional affiliations, we did not filter papers by geopolitical
region, and thus, a manuscript could rely solely on institutions within
a single country. The Physical Sciences did not show an increase in
diversity of regions represented, however, it could be argued that the
development of large research centers like CERN forced earlier diversity
than other sciences (i.e., because CERN specifically recruited
scientists from sponsoring nations). The Life, Health, and Social
Sciences saw an increase in the number of regions represented with the
highest increase in the Social Sciences. This result likely corresponds
with an increased interest in big team science publications in
psychology\textsuperscript{2,3}, and the desire to diversify the
populations represented in psychological research\textsuperscript{43,44}.

While publications on the whole are diversifying, we did not yet find
equality in the representation for first/corresponding author spots
versus all other authors. In general, first authors appear to be less
diverse, representing more European and North American authors, while
other authors include more Asian and African authors. These effect sizes
were often small, but the inequality still persists across years, even
if they are slowly decreasing. Diverse teams are more likely to have
papers with stronger ``impact''\textsuperscript{50,60--62} with higher citation metrics for more diverse author lists.
The introduction of contributorship models (e.g., CRediT\textsuperscript{42})
will hopefully continue to push these effects down, as they highlight
each individual's contribution to a manuscript.

@@

\leavevmode\vadjust pre{\hypertarget{ref-jones2008}{}}%
\CSLLeftMargin{50. }%
\CSLRightInline{Jones, B. F., Wuchty, S. \& Uzzi, B. \href{https://doi.org/10.1126/science.1158357}{Multi-university research teams: Shifting impact, geography, and stratification in science}. \emph{Science} \textbf{322}, 1259--1262 (2008).}

\leavevmode\vadjust pre{\hypertarget{ref-sinatra2015}{}}%
\CSLLeftMargin{51. }%
\CSLRightInline{Sinatra, R., Deville, P., Szell, M., Wang, D. \& Barabási, A.-L. \href{https://doi.org/10.1038/nphys3494}{A century of physics}. \emph{Nature Physics} \textbf{11}, 791--796 (2015).}

\leavevmode\vadjust pre{\hypertarget{ref-wuchty2007}{}}%
\CSLLeftMargin{52. }%
\CSLRightInline{Wuchty, S., Jones, B. F. \& Uzzi, B. \href{https://doi.org/10.1126/science.1136099}{The increasing dominance of teams in production of knowledge}. \emph{Science} \textbf{316}, 1036--1039 (2007).}

\leavevmode\vadjust pre{\hypertarget{ref-hunter2008}{}}%
\CSLLeftMargin{53. }%
\CSLRightInline{Hunter, L. \& Leahey, E. \href{https://doi.org/10.1007/s12108-008-9042-1}{Collaborative Research in Sociology: Trends and Contributing Factors}. \emph{The American Sociologist} \textbf{39}, 290--306 (2008).}

\leavevmode\vadjust pre{\hypertarget{ref-lariviuxe8re2015}{}}%
\CSLLeftMargin{54. }%
\CSLRightInline{Larivière, V., Gingras, Y., Sugimoto, C. R. \& Tsou, A. \href{https://doi.org/10.1002/asi.23266}{Team size matters: Collaboration and scientific impact since 1900}. \emph{Journal of the Association for Information Science and Technology} \textbf{66}, 1323--1332 (2015).}

\leavevmode\vadjust pre{\hypertarget{ref-lu2007}{}}%
\CSLLeftMargin{55. }%
\CSLRightInline{Lu, Y. \href{https://doi.org/10.1016/j.lisr.2006.10.007}{The human in human information acquisition: Understanding gatekeeping and proposing new directions in scholarship}. \emph{Library \& Information Science Research} \textbf{29}, 103--123 (2007).}

\leavevmode\vadjust pre{\hypertarget{ref-siler2015}{}}%
\CSLLeftMargin{56. }%
\CSLRightInline{Siler, K., Lee, K. \& Bero, L. \href{https://doi.org/10.1073/pnas.1418218112}{Measuring the effectiveness of scientific gatekeeping}. \emph{Proceedings of the National Academy of Sciences} \textbf{112}, 360--365 (2015).}

\leavevmode\vadjust pre{\hypertarget{ref-caplow2017}{}}%
\CSLLeftMargin{57. }%
\CSLRightInline{Caplow, T. \emph{The academic marketplace}. (Routledge, 2017). doi:\href{https://doi.org/10.4324/9781351305969}{10.4324/9781351305969}.}

\leavevmode\vadjust pre{\hypertarget{ref-kyvik2003}{}}%
\CSLLeftMargin{58. }%
\CSLRightInline{Kyvik, S. \href{https://doi.org/10.1023/A:1025475423482}{Changing trends in publishing behaviour among university faculty, 1980-2000}. \emph{Scientometrics} \textbf{58}, 35--48 (2003).}

\leavevmode\vadjust pre{\hypertarget{ref-xie2014}{}}%
\CSLLeftMargin{59. }%
\CSLRightInline{Xie, Y. \href{https://doi.org/10.1126/science.1252743}{{``}Undemocracy{''}: Inequalities in science}. \emph{Science} \textbf{344}, 809--810 (2014).}

\leavevmode\vadjust pre{\hypertarget{ref-freeman2015}{}}%
\CSLLeftMargin{60. }%
\CSLRightInline{Freeman, R. B. \& Huang, W. \href{https://doi.org/10.1086/678973}{Collaborating with People Like Me: Ethnic Coauthorship within the United States}. \emph{Journal of Labor Economics} \textbf{33}, S289--S318 (2015).}

\leavevmode\vadjust pre{\hypertarget{ref-hinnant2012}{}}%
\CSLLeftMargin{61. }%
\CSLRightInline{Hinnant, C. C. \emph{et al.} \href{https://doi.org/10.1016/j.lisr.2012.03.001}{Author-team diversity and the impact of scientific publications: Evidence from physics research at a national science lab}. \emph{Library \& Information Science Research} \textbf{34}, 249--257 (2012).}
